\chapter{Requirements}
The following sections list the requirements and their respective priorities.
In addition, it is indicated whether and how they are currently implemented.

\section{Non-Functional Requirements}

\subsection{Transparency (Must Have High)}\label{sec:Transparency}
\emph{Varroa has to be comprehensible for the user.}

This requirement is fulfilled.
Varroa implements expressive logging with configurable levels and outputs.
This enables rich insight in the execution process of a Varroa test.

\subsection{10.000.000 MQTT Clients (Must Have High)} 
\emph{Varroa has to be able to generate a large amount of clients.}

The fulfilment of this requirement is not verified yet.

\subsection{Scalability (Must Have)} 
\emph{Varroa should scale vertically with relatively low scaling costs.}

This requirement is fulfilled.
To do so Varroa is organised as a distributed System and can easily be extended with new instances running as Varroa Agents.

\subsection{Determinism (Must Have)} 
\emph{Varroa has to work in deterministic ways, meaning it should produce the same result for a Scenario every time.}

This requirement is fulfilled.
To ensure this determinism several measures are taken in the distribution and execution of a Scenario (see chapter \ref{sec:Architecture}).

\subsection{Distributed (Must Have Low)} 
\emph{Varroa is a distributed System.}

This requirement is fulfilled.
Varroa is organized as a distributed System, it is composed of a Commander and multiple Agents that can run on different machines. 
 
\subsection{Usabillity (Very Important)} 
\emph{Varroa has to be easily usable.}

This requirement is fulfilled.
The user can easily define custom Scenarios as XML files and execute them (see chapter \ref{sec:ScenarioConcept} and \ref{sec:Execution}).

\subsection{Code Quality (Important)} 
\emph{Varroa's code quality should be very high.}

This requirement is fulfilled.
To ensure high code quality several measures are taken.
For example the consequent use of nullability annotiations and a broad test code coverage.
Also before new code could be added to the master branch it was reviewed by another team member.

\subsection{Stability (Important)} 
\emph{Varroa has to run in a stable manner.}

This requirement is fulfilled.
By the use of integration tests and general unit tests stability is ensured.

\subsection{Resource efficiency (Important)} 
\emph{Varroa has to use the available computation and memory resources efficiently.}
%TODO how was this verified.

\subsection{User / Developer Guide (Somewhat Important)} 
\emph{Varroa needs a User / Developer Guide.}

This requirement is fulfilled. This documentation educates the user on defining a custom scenario and after that executing it on his Varroa Distributed System (\ref{sec:ScenarioConcept} and \ref{sec:Execution}).

\subsection{Automation capacity (Somewhat Important)} 
\emph{Varroa should be automatable.}

%TODO Docker.
%TODO is this fulfilled? 

\subsection{User Interface}
\emph{Varroa should have a well designed user interface with high usability.}

This requirement is fulfilled.
The user can interact with Varroa using a command line interface.

\subsection{Multi Platform Support (Nice To Have)}
\emph{Varroa should run on multiple platforms.}

This requirement is fulfilled.
Due to the utilisation of Docker, Varroa can run on several platforms (see chapter \ref{sec:Execution}).

\section{Functional Requirements}
The following sections list the non-functional requirements and their respective priorities.

\subsection{Scenarios (Must Have High)}
\emph{Varroa must be able to execute user defined MQTT Scenarios.}

This requirement is fulfilled.
Varroa has the ability to parse and execute Scenarios, according to the concept explained in chapter \ref{sec:ScenarioConcept}.

\subsection{MQTT Specification Confirmity (Must Have High)}
\emph{Varroa must conform to the MQTT specification.}

This requirement is fulfilled.
Varroa uses the HiveMQ MQTT Client library for its MQTT actions.
The library ensures specification conformity in both MQTT version 3 and MQTT version 5.

\subsection{All Transports (Must Have)}
\emph{Varroa has to support all transports that are possible in MQTT.}

This requirement is not fulfilled.
Currently Varroa does not support websockets.

\subsection{Reports (Must Have)}
\emph{Varroa has to report the findings of the testing process.}

This requirement is fulfilled.
Varroa supports reporting findings of tests (see \ref{sec:Reporting})

\subsection{User Feedback (Must Have Low)}
\emph{Varroa has to give user feedback during runtime.}

This requirement is fulfilled (see \ref{sec:Transparency}).

\subsection{Documentation (Very Important)}
\emph{Varroa has to be documented. }

This requirement is fulfilled.
This document contains a user guide and briefly explains the architecture and inner workings of Varroa.

\subsection{Metrics (Very Important)}
\emph{Varroa has to record and expose internal metrics.}

This requirement is not fulfilled.

\subsection{Deployment Convenience (Important)}
\emph{Varroa must be easily deployable.}

This requirement is fulfilled. Due to the use of Docker Varroa Instances can be easily started and so a Varroa Distributed System can be orchestrated without much configuration (see chapter \ref{sec:Execution}).

\subsection{Topology Change (Less important)}
\emph{Varroa must be able to simulate changes to the network topology.}

This requirement is not fulfilled.
It is not possible to simulate network level behaviour of the clients in the current version.

\subsection{Simple and Expert Mode (Somewhat Important}
\emph{Varroa has to have two user modes.
One for experienced users and one for unexperienced users.}

This requirement is not fulfilled.
Since Varroa does not have a graphical user interface in this version.
Simple and Expert modes are not implemented.


\subsection{Notifications (Nice To Have)}
\emph{Varroa must notitfy the user about the testing process by Email.}

This requirement is not fulfilled in the current version.


















